Viscous Transport in Eroding Porous Media

### 2019-08-21 docs/transport/erosion_transport.tex
small changes
replaced diffusion with dispersion in most locations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Flow in many natural occurring geometries involves flows that vary over
many orders of magnitude.  For example, the length scales of a porous
media can vary from $10^{-6}m$ to
$10^{-1}m$~\cite{mil-chr-imh-mcb-ped1998}.  Porous media flow plays an
important role in several environmental and industrial applications.
Therefore, understanding the dynamics of pore-scale flow is central to
upscale the governing fluid equations, characterize
dispersion~\cite{saf1959}, and quantify mixing. However, because of the
large disparity of length and velocity scales, constitutive laws that
link the microscopic scale to macroscopic scale must be
developed~\cite{mil-chr-imh-mcb-ped1998}.  Therefore, to find closure
formulas, accurate simulations at microscopic scales must be performed
and analyzed.  A few examples of closure formulas and coarse-grained
variables include
pressure-saturation-permeability~\cite{mil-chr-imh-mcb-ped1998}, first
passage time distribution~\cite{ber-sch-sil2000, hym-den-hag-kan2019,
cve-che-wen1996}, tortuosity~\cite{hak-com-den2019,
matyka2008tortuosity, dud-koz-mat2011, kop-kat-tim1996}, concentration
distributions~\cite{ica-den2019, bel-sal-rin1992}, flow
distributions~\cite{ali-par-wei-bre2017}, geometry
connectivity~\cite{knu-car2005, wes-blo-gra2001}, pore
distributions~\cite{ali-par-wei-bre2017}, and anomalous
diffusion~\cite{dea-qua-bir-jua2018, den-cor-sch-ber2004,
sie-ili-pri-riv-gua2019}.

In many applications, the fluid transport continuously changes the pore
structure.  This is particularly common in geophysical flows where
erosion alters the porous matrix.  In this work, we investigate flow in
eroded porous media.  By introducing erosion, there is a two-way
coupling between the hydrodynamics and the mechanical alteration of the
complex porous geometry.  Transport in porous geometries that have
undergone erosion finds applications in aquifers~\cite{cve-che-wen1996,
sud1986}, solute transport in groundwater flow~\cite{dag1987,
kon-bre1978}.  Porous geometries that have undergone erosion have
certain characteristics.  For example, the geometry often has regions of
high porosity that connect the inlet to the outlet.  This channelization
results in velocity scales that can vary over several orders of
magnitude~\cite{}.  Moreover, the channelization results in a geometry
that is largely heterogeneous and anisotropic.  Others have studied
transport in heterogeneous geometries (for example,
see~\cite{mil-chr-imh-mcb-ped1998, ber-sch-sil2000, hak-com-den2019,
suo-liu-gan2019, cus-hu-den1995, cve-che-wen1996, leb-ded-dav-bou2007,
knu-car2005, bel-sal-rin1992}). However, to the best of our knowledge, a
detailed study of transport in eroded geometries has not been performed.

To describe the relevant hydrodynamics, we model two-dimensional
microscopic scales where the fluid is governed by the incompressible
Stokes equations (for example, see~\cite{hym-den-hag-kan2019,
den-ica-hid2018, bar-mar-vee-zha2018}).  To couple the erosion and the
hydrodynamics, the grains of the porous media are eroded at a rate that
is proportional to the hydrodynamic shear stress~\cite{wan-fel2004,
ris-moo-chi-she-zha2012, qua-moo2018, par-izu2000}.  The incompressible
Stokes equation are solved by first recasting them as a boundary
integral equation (BIE).  The BIE is discretized with a high-order
quadrature method that resolves nearly touching bodies with a
near-singular integration scheme~\cite{bar-wu-vee2015}.  The erosion
dynamics are simulated by applying a stable second-order time stepping
Runge-Kutta method to the {\thL} formulation~\cite{hou-low-she1994} of
the eroding bodies.  Given the solution of the incompressible Stokes
equations are solved in an eroded porous geometry, we investigate
different metrics to characterize transport.  We focus on three metrics:
the tortuosity, the anomalous diffusion, and the distribution of
geometrical channels.

The first two of these metrics can be defined in terms of statistical
moments of trajectories.  Depending on the application, transport at the
microscale can be modelled as pure advection~\cite{dea-qua-bir-jua2018,
leb-ded-dav-bou2007, cve-che-wen1996, puy-gou-den2019}, and advection
diffusion equation~\cite{cus-hu-den1995, dag1987, den-ica-hid2018}, or
with a random walk~\cite{saf1959, bij-blu2006, ber-sch-sil2000}.  In
this work, we assume that the trajectories are governed purely by
advection (ie.~no diffusion).  Therefore, a trajectory $\ss(t)$
initialized at $\ss_0$ is governed by
\begin{align}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0,
\end{align}
where $\uu$ is the fluid velocity.  Since there is no diffusion, all
dispersion is a result of the particle spreading induced by the porous
geometry.

@@

Anomalous diffusion  ...
\cite{cus-hu-den1995}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions}
\begin{itemize}
  \item Use novel near-singular Barycentric integration scheme for
    Stokes equation outlined in~\cite{bar-wu-vee2015}. (Remember to
    include historically significant reference Shelley
    \cite{baker1986boundary}).  Also need discuss Helsing and Ojala who
    did a similar panel-based quadrature~\cite{hel-oja2008}
  \item Need to compute one additional derivative since we need the
    derivative of the velocity to find the vorticity and stress tensor
  \item Annealing algorithm to create initial body configurations of specified area fraction specified size distribution (NM can do the writing for this).
  \item Tracer trajectories at different time steps
  \item Compute the anomalous diffusion rate using similar techniques as
    the paper with Pietro~\cite{dea-qua-bir-jua2018}
  \item Need a novel way to do time stepping through the complex
    geometry
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport, Tracers, and Tortuosity}
\label{sec:transport}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Erosion leads to interesting porous geometries, and we are interested in
characterizing transport in these geometries.  For porous media, two of
the most important variables are the porosity and the permeability.
However, additional properties of the flow are revealed by computing
the anomalous diffusion rate, the tortuosity, and the distribution the
gap sizes.

Having developed numerically accurate methods to compute the velocity
$\uu$ at points close to the bodies, we are able to compute the
trajectories of particles passing through an eroded body.  Since the
trajectories are computed with high-order methods in both space and
time,  we are able to simulate dynamics very close to the eroding bodies
(see Figures~\ref{fig:Eroding20tracer} and~\ref{fig:Eroding100tort}.
Resolving the flow in these small velocity regions is essential to
accurately describe anomalous diffusion and
mixing~\cite{leb-den-dav-bol-car-dec-bou2011, den-leb-eng-bij2011}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Anomalous Diffusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The dynamics of fluid in porous media is often characterized in terms of
anomalous diffusion~\cite{kla-rad-sok2008, den-cor-sch-ber2004}.  The
anomalous diffusion depends on the porosity and
permeability~\cite{koc-bra1988}, but the distribution of the bodies also
affects the anomalous diffusion rates.  Since erosion causes certain
structure in porous media, calculating the anomalous diffusion rates in
these geometries is important.

We calculate the anomalous diffusion rates by analyzing the tracer
trajectories governed by equation~\eqref{eqn:trajectories} for
geometries that are formed by erosion.  Given a set of trajectories, we
define $\lambda_j(t)$ to be the arclength of the trajectory
\begin{align}
  \lambda_j(t) = \int_{0}^t \|\ss'_j(\tau)\|\, d\tau, 
    \quad j=1,\ldots,N_p.
\end{align}
Then, the first and second ensemble moments are
\begin{align}
  \langle \lambda \rangle (t) = 
    \frac{1}{N_p} \sum_{j=1}^{N_p} \lambda_j(t), \quad 
    \sigma_\lambda^2(t) = \frac{1}{N_p} 
    \left[\lambda_j(t) - \langle \lambda \rangle(t) \right]^2.
\end{align}
Since we compute second-order statistics, we use 1000 tracers so that
the statistics have converged~\cite{bel-sal-rin1992}.  Then, particle
dispersion is characterized as $\sqrt{\sigma_\lambda^2}$.  At early
times, the particles have not explored much of the geometry, and we
expect a ballistic motion $\sigma_\lambda^2 \sim t^2$.  However, as the
particles pass the grains, their trajectories are altered, and we expect
that $\sigma_\lambda \sim t^\alpha$, with $\alpha < 1$, indicating that
the flow is sub-diffusive.  With respect to erosion, as the bodies
erode, we expect that the channelization of the geometry results in more
trajectories undergoing a ballistic motion, and this would indicate
decrease particle dispersion, meaning a larger anomalous diffusion rate.

@@

  \item{\bf 20 Bodies at a Medium Porosity}: We consider 20 eroding
  bodies with a medium initial porosity.  After computing accurate
  tracer trajectories, the tortuosity and anomalous diffusion rates are
  computed and compared to those of an open channel.

  \item{\bf 20 Bodies at a Low Porosity}: We consider 20 eroding bodies
  with a low initial porosity.  We examine the effect of the lower
  porosity on the tortuosity and anomalous diffusion rates.

  \item{\bf 50 Bodies at a Low Porosity}: We consider 50 eroding bodies
  with a low initial porosity.  In addition to the tortuosity and
  anomalous diffusion rates, we compute distributions of the pore throat
  sizes and examine their response to erosion.

  \item{\bf 100 Bodies at a Medium Porosity}: We consider 100 eroding
  bodies with a medium porosity.  We compute the tortuosity, anomalous
  diffusion rates, and the pore throat distributions.
\end{itemize}

@@

Next, we use the tracer trajectories to compute the anomalous diffusion
rates and the tortuosities in the eroding geometry.  Since these
calculations use a statistical analysis of the tracer trajectories, we
increase the number of trajectories to $N_p = 1000$.  Again, the tracers
are initialized at $x=-1$, and their $y$ coordinates are distributed
evenly across the entire channel.  To compute the tortuosity, we
require the velocity at the initialization points of the tracers.  These
normalized velocities are plotted in Figure~ref{fig:Eroding20tort}(a)
for the eroded geometry with porosity 62.9\% (see
Figure~\ref{fig:Eroding20tort}(c)).  The velocities are similar to the
work of Matyka et al.~\cite{matyka2008tortuosity} (see Figure 4(a)),
except that our cross-section does not cut through any of the bodies.
Next, in Figure~\ref{fig:Eroding20tort}(b), we plot the local
tortuosity~\eqref{eqn:localTort} by calculating the length of each
trajectory as it traverses the channel from $x=-1$ to $x=1$ in
Figure~\ref{fig:Eroding20tort}(b), and dividing by the length of the
channel.  The local tortuosity ranges from 1 to 1.27, meaning that one
of the tracers travelled 27\% farther than it would have if the bodies
had been absent.  The average tracer travelled 9.79\% farther than if
the bodies had been absent, meaning that the tortuosity of the eroded
channel is $1.098$.  Again, comparing the local tortuosity to
Matyka~\cite{matyka2008tortuosity} (Figure 4(b)), the results are
qualitatively similar. However, since our initial cross-section does not
cut through any bodies, the local tortuosity does not have any gaps.
The discontinuities correspond to neighboring tracers with one traveling
much farther than the other.  In Figure~\ref{fig:Eroding20tort}(c), we
plot pairs of trajectories that correspond to the ten largest jumps in
the local tortuosity.  Corresponding pairs are always plotted in the
same color.  The discrepancy between the trajectory lengths can be
attributed to the tracers diverging from one another as they approach
opposite sides of a stagnation point in the flow.

@@

\begin{figure}[H]
%\caption{\label{fig:Eroding20anomalous} The
%mean squared displacement of the tracer trajectories  at several
%different stages of the erosion process.  As the porosity increases, the
%mean squared displacement tends towards $\sigma^2_\lambda(t) \sim t^$
%}
%\includegraphics*[width =0.45\linewidth]{./figs/20b_first_moment}
\center
\includegraphics*[width =0.45\linewidth]{./figs/20b_second_moment_long}
\caption{\label{fig:Eroding20anomalous} The analysis of anomalous
diffusion in 20 nearly touching bodies with respect to initial porosity (37.68\%) and another six porosities
during the eroding. The dashed line correspondes to a porosity of 100\% and the dashed-dotted line is a 
line of best fit with slope 0.5742.}
\end{figure}

@@

Next, we use the 1000 tracer trajectories to compute the first
(Figure~\ref{fig:Eroding100anomalous}(a)) and second
(Figure~\ref{fig:Eroding100anomalous}(b)) moments of the streamlines at
various porosities.  At high porosities, the first moment is similar to
the first-moment of trajectories in a empty channel (black dashed line).
At smaller porosities, the first moment is always linear, but larger
slopes indicate that the length of the trajectories are longer at lower
porosities \todo[inline]{Don't understand how this is the case}.  The
onset of anomalous diffusion is observed in
Figure~\ref{fig:Eroding100anomalous}(b). At early times, the particle
spreading is comparable to that of an open channel (dashed curve)
corresponding to the ballistic regime.  However, as the trajectories
traverse the geometry, a transition to a sub-diffusion case, especially
at low porosities, is evident.  The dashed-dotted line corresponds has a
slope less than 1 indicating the long time behavior of the trajectories
\begin{align*}
  \sigma_\lambda \sim t^{0.8715}.
\end{align*}

@@

Finally, using network models for flow in porous media~\cite{}, the
anomalous diffusion rate can be linked to channel widths between solid
bodies.  We construct a Delaunay triangulation with the centers of the
eroding bodies corresponding to the vertices of the triangles.  Then, if
the side of a triangle connects two vertices, we call the bodies
neighbors.  Then, for each pair of neighbors, we measure the distance
between the bodies.  Once a Delaunay triangulation is formed, it is used
for all subsequent time steps until a body completely disappears.  At
this point, a new Delaunay triangulation is formed.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Flow in many natural occurring geometries involves flows that vary over
many orders of magnitude.  For example, the length scales of a porous
media can vary from $10^{-6}m$ to
$10^{-1}m$~\cite{mil-chr-imh-mcb-ped1998}.  Porous media flow plays an
important role in several environmental and industrial applications.
Therefore, understanding the dynamics of pore-scale flow is central to
upscale the governing fluid equations, characterize
dispersion~\cite{saf1959}, and quantify mixing. However, because of the
large disparity of length and velocity scales, constitutive laws that
link the microscopic scale to macroscopic scale must be
developed~\cite{mil-chr-imh-mcb-ped1998}.  Therefore, to find closure
formulas, accurate simulations at microscopic scales must be performed
and analyzed.  A few examples of closure formulas and coarse-grained
variables include
pressure-saturation-permeability~\cite{mil-chr-imh-mcb-ped1998}, first
passage time distribution~\cite{ber-sch-sil2000, hym-den-hag-kan2019,
cve-che-wen1996}, tortuosity~\cite{hak-com-den2019,
matyka2008tortuosity, dud-koz-mat2011, kop-kat-tim1996}, concentration
distributions~\cite{ica-den2019, bel-sal-rin1992}, flow
distributions~\cite{ali-par-wei-bre2017}, geometry
connectivity~\cite{knu-car2005, wes-blo-gra2001}, pore
distributions~\cite{ali-par-wei-bre2017}, and anomalous
dispersion~\cite{dea-qua-bir-jua2018, den-cor-sch-ber2004,
sie-ili-pri-riv-gua2019}.

In many applications, the fluid transport continuously changes the pore
structure.  This is particularly common in geophysical flows where
erosion alters the porous matrix.  In this work, we investigate flow in
eroded porous media.  By introducing erosion, there is a two-way
coupling between the hydrodynamics and the mechanical alteration of the
complex porous geometry.  Transport in porous geometries that have
undergone erosion finds applications in aquifers~\cite{cve-che-wen1996,
sud1986}, solute transport in groundwater flow~\cite{dag1987,
kon-bre1978}.  Porous geometries that have undergone erosion have
certain characteristics.  For example, the geometry often has regions of
high porosity that connect the inlet to the outlet.  This channelization
results in velocity scales that can vary over several orders of
magnitude~\cite{all-hea-lab-rei2002}.  Moreover, the channelization
results in a geometry that is largely heterogeneous and anisotropic.
Others have studied transport in heterogeneous geometries (for example,
see~\cite{mil-chr-imh-mcb-ped1998, ber-sch-sil2000, hak-com-den2019,
suo-liu-gan2019, cus-hu-den1995, cve-che-wen1996, leb-ded-dav-bou2007,
knu-car2005, bel-sal-rin1992}). However, to the best of our knowledge, a
detailed study of transport in eroded geometries has not been performed.

To describe the relevant hydrodynamics, we model two-dimensional
microscopic scales where the fluid is governed by the incompressible
Stokes equations (for example, see~\cite{hym-den-hag-kan2019,
den-ica-hid2018, bar-mar-vee-zha2018}).  To couple the erosion and the
hydrodynamics, the grains of the porous media are eroded at a rate that
is proportional to the hydrodynamic shear stress~\cite{wan-fel2004,
ris-moo-chi-she-zha2012, qua-moo2018, par-izu2000}.  The incompressible
Stokes equation are solved by first recasting them as a boundary
integral equation (BIE).  The BIE is discretized with a high-order
quadrature method that resolves nearly touching bodies with a
near-singular integration scheme~\cite{bar-wu-vee2015}.  The erosion
dynamics are simulated by applying a stable second-order time stepping
Runge-Kutta method to the {\thL} formulation~\cite{hou-low-she1994} of
the eroding bodies.  Given the solution of the incompressible Stokes
equations are solved in an eroded porous geometry, we investigate
different metrics to characterize transport.  We focus on three metrics:
the tortuosity, the anomalous dispersion, and the distribution of
geometrical channels.

The first two of these metrics can be defined in terms of statistical
moments of trajectories.  Depending on the application, transport at the
microscale can be modelled as pure advection~\cite{dea-qua-bir-jua2018,
leb-ded-dav-bou2007, cve-che-wen1996, puy-gou-den2019}, and advection
dispersion equation~\cite{cus-hu-den1995, dag1987, den-ica-hid2018}, or
with a random walk~\cite{saf1959, bij-blu2006, ber-sch-sil2000}.  In
this work, we assume that the trajectories are governed purely by
advection (ie.~no diffusion).  Therefore, a trajectory $\ss(t)$
initialized at $\ss_0$ is governed by
\begin{align}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0,
\end{align}
where $\uu$ is the fluid velocity.  Since there is no diffusion, all
dispersion is a result of the particle spreading induced by the porous
geometry.

@@

The tortuosity is a first statistical moment of the tracer trajectories.
However, to measure spread of trajectories, second-order moments are
computed. By computing the variance of the lengths of trajectories in a
porous media, non-local anomalous dispersion is
observed~\cite{kan-dea-nun-bij-blu-jua2014, cus-hu-den1995,
dea-leb-den-tar-bol-dav2013}, resulting in super-diffusive dispersion.
Anomalous dispersion has been investigated for a variety of porous
geometries, but the effect of erosion on the anomalous dispersion rate
is lacking.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions}
\begin{itemize}
  \item Use novel near-singular Barycentric integration scheme for
    Stokes equation outlined in~\cite{bar-wu-vee2015}. (Remember to
    include historically significant reference Shelley
    \cite{baker1986boundary}).  Also need discuss Helsing and Ojala who
    did a similar panel-based quadrature~\cite{hel-oja2008}
  \item Need to compute one additional derivative since we need the
    derivative of the velocity to find the vorticity and stress tensor
  \item Annealing algorithm to create initial body configurations of specified area fraction specified size distribution (NM can do the writing for this).
  \item Tracer trajectories at different time steps
  \item Compute the anomalous dispersion rate using similar techniques
  as the paper with Pietro~\cite{dea-qua-bir-jua2018}
  \item Need a novel way to do time stepping through the complex
    geometry
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport, Tracers, and Tortuosity}
\label{sec:transport}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Erosion leads to interesting porous geometries, and we are interested in
characterizing transport in these geometries.  For porous media, two of
the most important variables are the porosity and the permeability.
However, additional properties of the flow are revealed by computing
the anomalous dispersion rate, the tortuosity, and the distribution the
gap sizes.

Having developed numerically accurate methods to compute the velocity
$\uu$ at points close to the bodies, we are able to compute the
trajectories of particles passing through an eroded body.  Since the
trajectories are computed with high-order methods in both space and
time,  we are able to simulate dynamics very close to the eroding bodies
(see Figures~\ref{fig:Eroding20tracer} and~\ref{fig:Eroding100tort}.
Resolving the flow in these small velocity regions is essential to
accurately describe anomalous dispersion and
mixing~\cite{leb-den-dav-bol-car-dec-bou2011, den-leb-eng-bij2011}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Anomalous Diffusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The dynamics of fluid in porous media is often characterized in terms of
anomalous dispersion~\cite{kla-rad-sok2008, den-cor-sch-ber2004}.  The
anomalous dispersion depends on the porosity and
permeability~\cite{koc-bra1988}, but the distribution of the bodies also
affects the anomalous dispersion rates.  Since erosion causes certain
structure in porous media, calculating the anomalous dispersion rates in
these geometries is important.

We calculate the anomalous dispersion rates by analyzing the tracer
trajectories governed by equation~\eqref{eqn:trajectories} for
geometries that are formed by erosion.  Given a set of trajectories, we
define $\lambda_j(t)$ to be the arclength of the trajectory
\begin{align}
  \lambda_j(t) = \int_{0}^t \|\ss'_j(\tau)\|\, d\tau, 
    \quad j=1,\ldots,N_p.
\end{align}
Then, the first and second ensemble moments are
\begin{align}
  \langle \lambda \rangle (t) = 
    \frac{1}{N_p} \sum_{j=1}^{N_p} \lambda_j(t), \quad 
    \sigma_\lambda^2(t) = \frac{1}{N_p} 
    \left[\lambda_j(t) - \langle \lambda \rangle(t) \right]^2.
\end{align}
Since we compute second-order statistics, we use 1000 tracers so that
the statistics have converged~\cite{bel-sal-rin1992}.  Then, particle
dispersion is characterized as $\sqrt{\sigma_\lambda^2}$.  At early
times, the particles have not explored much of the geometry, and we
expect a ballistic motion $\sigma_\lambda^2 \sim t^2$.  However, as the
particles pass the grains, their trajectories are altered, and we expect
that $\sigma_\lambda \sim t^\alpha$, with $\alpha < 1$, indicating that
the flow is sub-diffusive.  With respect to erosion, as the bodies
erode, we expect that the channelization of the geometry results in more
trajectories undergoing a ballistic motion, and this would indicate
decrease particle dispersion, meaning a larger anomalous dispersion
rate.

@@

  \item{\bf 20 Bodies at a Medium Porosity}: We consider 20 eroding
  bodies with a medium initial porosity.  After computing accurate
  tracer trajectories, the tortuosity and anomalous dispersion rates are
  computed and compared to those of an open channel.

  \item{\bf 20 Bodies at a Low Porosity}: We consider 20 eroding bodies
  with a low initial porosity.  We examine the effect of the lower
  porosity on the tortuosity and anomalous dispersion rates.

  \item{\bf 50 Bodies at a Low Porosity}: We consider 50 eroding bodies
  with a low initial porosity.  In addition to the tortuosity and
  anomalous dispersion rates, we compute distributions of the pore
  throat sizes and examine their response to erosion.

  \item{\bf 100 Bodies at a Medium Porosity}: We consider 100 eroding
  bodies with a medium porosity.  We compute the tortuosity, anomalous
  dispersion rates, and the pore throat distributions.
\end{itemize}

@@

Next, we use the tracer trajectories to compute the anomalous dispersion
rates and the tortuosities in the eroding geometry.  Since these
calculations use a statistical analysis of the tracer trajectories, we
increase the number of trajectories to $N_p = 1000$.  Again, the tracers
are initialized at $x=-1$, and their $y$ coordinates are distributed
evenly across the entire channel.  To compute the tortuosity, we
require the velocity at the initialization points of the tracers.  These
normalized velocities are plotted in Figure~ref{fig:Eroding20tort}(a)
for the eroded geometry with porosity 62.9\% (see
Figure~\ref{fig:Eroding20tort}(c)).  The velocities are similar to the
work of Matyka et al.~\cite{matyka2008tortuosity} (see Figure 4(a)),
except that our cross-section does not cut through any of the bodies.
Next, in Figure~\ref{fig:Eroding20tort}(b), we plot the local
tortuosity~\eqref{eqn:localTort} by calculating the length of each
trajectory as it traverses the channel from $x=-1$ to $x=1$ in
Figure~\ref{fig:Eroding20tort}(b), and dividing by the length of the
channel.  The local tortuosity ranges from 1 to 1.27, meaning that one
of the tracers travelled 27\% farther than it would have if the bodies
had been absent.  The average tracer travelled 9.79\% farther than if
the bodies had been absent, meaning that the tortuosity of the eroded
channel is $1.098$.  Again, comparing the local tortuosity to
Matyka~\cite{matyka2008tortuosity} (Figure 4(b)), the results are
qualitatively similar. However, since our initial cross-section does not
cut through any bodies, the local tortuosity does not have any gaps.
The discontinuities correspond to neighboring tracers with one traveling
much farther than the other.  In Figure~\ref{fig:Eroding20tort}(c), we
plot pairs of trajectories that correspond to the ten largest jumps in
the local tortuosity.  Corresponding pairs are always plotted in the
same color.  The discrepancy between the trajectory lengths can be
attributed to the tracers diverging from one another as they approach
opposite sides of a stagnation point in the flow.

@@

\begin{figure}[H]
%\caption{\label{fig:Eroding20anomalous} The
%mean squared displacement of the tracer trajectories  at several
%different stages of the erosion process.  As the porosity increases, the
%mean squared displacement tends towards $\sigma^2_\lambda(t) \sim t^$
%}
%\includegraphics*[width =0.45\linewidth]{./figs/20b_first_moment}
\center
\includegraphics*[width =0.45\linewidth]{./figs/20b_second_moment_long}
\caption{\label{fig:Eroding20anomalous} The analysis of anomalous
dispersion in 20 nearly touching bodies with respect to initial porosity
(37.68\%) and another six porosities during the eroding. The dashed line
correspondes to a porosity of 100\% and the dashed-dotted line is a line
of best fit with slope 0.5742.}
\end{figure}

@@

Next, we use the 1000 tracer trajectories to compute the first
(Figure~\ref{fig:Eroding100anomalous}(a)) and second
(Figure~\ref{fig:Eroding100anomalous}(b)) moments of the streamlines at
various porosities.  At high porosities, the first moment is similar to
the first-moment of trajectories in a empty channel (black dashed line).
At smaller porosities, the first moment is always linear, but larger
slopes indicate that the length of the trajectories are longer at lower
porosities \todo[inline]{Don't understand how this is the case}.  The
onset of anomalous dispersion is observed in
Figure~\ref{fig:Eroding100anomalous}(b). At early times, the particle
spreading is comparable to that of an open channel (dashed curve)
corresponding to the ballistic regime.  However, as the trajectories
traverse the geometry, a transition to a super-dispersive case,
especially at low porosities, is evident.  The dashed-dotted line
corresponds has a slope less than 1 indicating the long time behavior of
the trajectories
\begin{align*}
  \sigma_\lambda \sim t^{0.8715}.
\end{align*}

@@

Finally, using network models for flow in porous media~\cite{}, the
anomalous dispersion rate can be linked to channel widths between solid
bodies.  We construct a Delaunay triangulation with the centers of the
eroding bodies corresponding to the vertices of the triangles.  Then, if
the side of a triangle connects two vertices, we call the bodies
neighbors.  Then, for each pair of neighbors, we measure the distance
between the bodies.  Once a Delaunay triangulation is formed, it is used
for all subsequent time steps until a body completely disappears.  At
this point, a new Delaunay triangulation is formed.
